# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
Let's read the zipped csv file and omit rows without steps.
```{r}
data <- read.csv(unz("activity.zip", "activity.csv"))
completeRows <- na.omit(data)
```

## What is mean total number of steps taken per day?

First, let's group by date:
```{r}
stepsPerDay <- aggregate(completeRows$steps, by=list(completeRows$date), FUN=sum)
```
Histogram of total number of steps taken per day looks like this:
```{r}
hist(stepsPerDay$x, breaks = nrow(stepsPerDay), main = "Histogram of total number of steps per day", xlab="Steps per day")
```
Then, let's calculate the mean and median:
```{r}
meanStepsPerDay = mean(stepsPerDay$x)
medianStepsPerDay = median(stepsPerDay$x)
```

Mean total number of steps taken per day is `r meanStepsPerDay`.
Median total number of steps taken per day is `r medianStepsPerDay`.

## What is the average daily activity pattern?

First, let's group by interval:
```{r}
averageStepsPerInterval <- aggregate(completeRows$steps, by=list(completeRows$interval), FUN=mean)
```
Then we can draw a time series plot:
```{r}
require(ggplot2)
ggplot(data = averageStepsPerInterval, aes(Group.1, x)) + 
  geom_line() + 
  xlab("Interval") + 
  ylab("Average number of steps")
````
Then let's fetch maximum number of steps:
```{r}
maxAverageStepsPerInterval <- averageStepsPerInterval[which.max(averageStepsPerInterval$x),]
```

On average, maximum number of steps per interval is `r maxAverageStepsPerInterval[[1, 2]]`, which is on interval `r maxAverageStepsPerInterval[[1, 1]]`.


## Imputing missing values

Number of rows with missing values is the difference between all rows and complete rows:
```{r}
numberOfRowsWithMissingValues <- nrow(data) - nrow(completeRows)
```

Number of rows with missing values is `r numberOfRowsWithMissingValues`.

Let's create a new data frame with NAs replaced with mean steps for the given five minute interval:
```{r}
naRows <- data[is.na(data$steps), ]
replacedRows <- NULL
for (rowIndex in 1:nrow(naRows)) {
  interval <- naRows[[rowIndex,3]]
  replacedSteps <- averageStepsPerInterval[averageStepsPerInterval$Group.1 == interval,][[1, 2]]
  replacedRows <- rbind(replacedRows, data.frame("steps" = replacedSteps, "date" = naRows[[rowIndex,2]], "interval" = interval))
}
```

Then let's append complete rows and these new replaced rows. Also, let's group by date:
```{r}
missingValuesFilled <- rbind(completeRows, replacedRows)
stepsPerDayFilled <- aggregate(missingValuesFilled$steps, by=list(missingValuesFilled$date), FUN=sum)
```

Histogram of the new data set looks like this:
```{r}
hist(stepsPerDayFilled$x, breaks = nrow(stepsPerDayFilled), main="Histogram of total number of steps per day with filled data set", xlab="Steps per day")
```
Like before, let's calculate mean and median of from filled data:
```{r}
meanStepsPerDayFilled = mean(stepsPerDayFilled$x)
medianStepsPerDayFilled = median(stepsPerDayFilled$x)
```

And the mean and median values are `r meanStepsPerDayFilled` and `r medianStepsPerDayFilled`.

So, if we did not make any mistakes, filling the steps modifies the median to be the same as mean.

## Are there differences in activity patterns between weekdays and weekends?

Let's split data by interval and weekday/weekend:
```{r}
isWeekend <- (weekdays(as.Date(missingValuesFilled$date)) %in% c("Saturday", "Sunday"))
stepsPerInterval <- aggregate(missingValuesFilled$steps, by=list(missingValuesFilled$interval, isWeekend), FUN=mean)
```

Now we can plot both groups:
```{r}
require("lattice")
factor <- factor(stepsPerInterval$Group.2, labels = c("Weekday", "Weekend"))
xyplot(stepsPerInterval$x ~ stepsPerInterval$Group.1 | factor, type = "l", xlab = "Interval", ylab = "Average number of steps")
```
